$Z_3$ topological order in the quantum dimer-pentamer model

### 2016-11-27 qdpm_paper.tex
v primes in appendix fixed
included correlation lengths

%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{spatial_cors_parallel.pdf}
    \caption{The dimer-dimer density correlation function $C_d$ between parallel links in the QDPM RK ground state on a $L=64$ lattice, in directions perpendicular (circles) and parallel (squares) to the dimer orientation.  The lines represent the best fit to an exponential decay with correlation length $\chi\approx???$. \note{Owen: Can you add the direction parallel to the dimers to this plot? Can you add the correlation length estimate here as well?}}
    \label{fig:spatial_dmr_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

To investigate the $Z_3$ gauge theory description of the QDPM RK ground state, we study the $Z_3$ magnetic string correlation function as defined in \eqref{eq:Sgamma} \& \eqref{eq:SQDPM}:
\begin{equation}
C_M \left( \bm{r}_{pp'} \right) \equiv \left \langle \exp \left[ \frac{2 \pi}{3} i \left( \Phi_{\gamma_{pp'}} -\tilde{\Phi}_{\gamma_{pp'}} \right) \right] \right \rangle
\end{equation}
where $p$ and $p'$ are two plaquettes, and $\gamma_{pp'}$ is any open string on the dual lattice connecting $p$ and $p'$. In \figref{fig:dimer_vison_heatmap}, we have plotted $|C_M|$ for an $L=16$ lattice. Note, that while $C_M$ depends on the choice of the background dimerization which fixes the gauge, $|C_M|$ is independent of the background dimerization. Additionally, we have computed the magnetic string structure factor
\begin{equation}
S_M\left( \bm{k} \right) \equiv \frac{1}{ L^2} \sum_{p,p'} e^{-i \bm{k} \cdot r_{pp'}} C_M\left( \bm{r}_{pp'}\right)
\end{equation}
and plot the magnitude $|S_M$ in \figref{fig:structure_factor}. We have chosen d a columnar background dimerization comprising horizontal background dimers stacked in columns along the $y$-axis, as shown in \figref{fig:BKGD_dimers}\note{Where should we show the background dimerization??}; this chose breaks the translational symmetry and thus requires doubling the unit cell the $x$-direction. \note{Owen: What are the effects of the background dimerization?}. While $|C_M|$ displays no obvious symmetry breaking, there are peaks in $|S_M|$ at $(0,0)$ \& $(0,\pi$) \note{(Owen: What about $(\pi,\pi)$??)}. We have plotted the finite-size scaling of these peaks in \figref{fig:bragg_scaling} which demonstrates that these peaks vanish in the thermodynamic limit. While the vanishing of the $(0,\pi)$ peak demonstrates the restoration of rotational and translational symmetries in the thermodynamic limit, the vanishing of the $(0,0)$ peak demonstrates the lack of magnetic vortex condensation, which is a mechanism for destroying topological order and breaking lattice symmetries~\cite{Jalabert1991,Ralko2007,Huh2011}. \figref{fig:vison_cor} shows the decay of $|C_M|$ along one of the axes of the dual lattice; \note{we find this decays exponentially with distance, with a correlation length $\chi\approx???$}. These exponentially decaying magnetic string correlations are therefore consistent with a symmetric dimer liquid with no magnetic vortex condensation.

@@

Here we will review a string-net implementation of $Z_3$ gauge theory on the square lattice~\cite{Horn1979}. We define a three -dimensional Hilbert space on each link, corresponding to the presence or absence of an oriented string segment that carries $Z_3$ electric flux which are eigenstates of the Hermitian $Z_3$ electric flux operator $E_{vv'}$
\begin{align}
        & E_{vv'}  \eket{\vvemptylink} = 0
        ,
        \\
        & E_{vv'}  \eket{\rightarrowlink} = + \eket{\rightarrowlink}
        ,
        \\
        & E_{vv'} \eket{\leftarrowlink} = - \eket{\leftarrowlink}
        .
\end{align}
\note{Owen:can you add symbols with $v$ and $v'$ at both end points for all types of links?}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{spatial_cors_parallel.pdf}
    \caption{The dimer-dimer density correlation function $C_d$ between parallel links in the QDPM RK ground state on a $L=64$ lattice, in directions perpendicular (circles) and parallel (squares) to the dimer orientation.  The lines represent the best fit to an exponential decay with correlation length $\chi\approx 1.1$.}
    \label{fig:spatial_dmr_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

To investigate the $Z_3$ gauge theory description of the QDPM RK ground state, we study the $Z_3$ magnetic string correlation function as defined in \eqref{eq:Sgamma} \& \eqref{eq:SQDPM}:
\begin{equation}
C_M \left( \bm{r}_{pp'} \right) \equiv \left \langle \exp \left[ \frac{2 \pi}{3} i \left( \Phi_{\gamma_{pp'}} -\tilde{\Phi}_{\gamma_{pp'}} \right) \right] \right \rangle
\end{equation}
where $p$ and $p'$ are two plaquettes, and $\gamma_{pp'}$ is any open string on the dual lattice connecting $p$ and $p'$. In \figref{fig:dimer_vison_heatmap}, we have plotted $|C_M|$ for an $L=16$ lattice. Note, that while $C_M$ depends on the choice of the background dimerization which fixes the gauge, $|C_M|$ is independent of the background dimerization. Additionally, we have computed the magnetic string structure factor
\begin{equation}
S_M\left( \bm{k} \right) \equiv \frac{1}{ L^2} \sum_{p,p'} e^{-i \bm{k} \cdot r_{pp'}} C_M\left( \bm{r}_{pp'}\right)
\end{equation}
and plot the magnitude $|S_M$ in \figref{fig:structure_factor}. We have chosen d a columnar background dimerization comprising horizontal background dimers stacked in columns along the $y$-axis, as shown in \figref{fig:BKGD_dimers}\note{Where should we show the background dimerization??}; this chose breaks the translational symmetry and thus requires doubling the unit cell the $x$-direction. \note{Owen: What are the effects of the background dimerization?}. While $|C_M|$ displays no obvious symmetry breaking, there are peaks in $|S_M|$ at $(0,0)$ \& $(0,\pi$) \note{(Owen: What about $(\pi,\pi)$??)}. We have plotted the finite-size scaling of these peaks in \figref{fig:bragg_scaling} which demonstrates that these peaks vanish in the thermodynamic limit.  While the vanishing of the $(0,\pi)$ peak demonstrates the restoration of rotational and translational symmetries in the thermodynamic limit, the vanishing of the $(0,0)$ peak demonstrates the lack of magnetic vortex condensation, which is a mechanism for destroying topological order and breaking lattice symmetries~\cite{Jalabert1991,Ralko2007,Huh2011}. \figref{fig:vison_cor} shows the decay of $|C_M|$ along one of the axes of the dual lattice; \note{we find this decays exponentially with distance, with a correlation length $\chi\approx 2.90$}. These exponentially decaying magnetic string correlations are therefore consistent with a symmetric dimer liquid with no magnetic vortex condensation.

@@

Here we will review a string-net implementation of $Z_3$ gauge theory on the square lattice~\cite{Horn1979}. We define a three -dimensional Hilbert space on each link, corresponding to the presence or absence of an oriented string segment that carries $Z_3$ electric flux which are eigenstates of the Hermitian $Z_3$ electric flux operator $E_{vv'}$
\begin{align}
        & E_{vv'}  \eket{\vvemptylink} = 0
        ,
        \\
        & E_{vv'}  \eket{\rightarrowlink} = + \eket{\rightarrowlink}
        ,
        \\
        & E_{vv'} \eket{\leftarrowlink} = - \eket{\leftarrowlink}
        .
\end{align}
